---
title: "Flux flux_selected introducing"
date: "02/02/2024"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

### Step 0: load your library flux_selectedset 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(corrplot)
library(psych)
# rm(list = ls())
```

### Step 1: Explore your flux_selectedset 
flux_selected cleanup, units conversion; typical ranges, annual sums and averages  

**Meteorological flux_selected:** 
TIMEKEEPING
TIMESTAMP - Time of the measurement (YEAR, MONTH, and DAY) ========= Time(TIMESTAMP)

MICROMETEOROLOGICAL
TA_F_MDS - Air temperature (°C) ================================     Air temperature(TA)
SW_IN_F - Incoming shortwave radiation, SWDOWN (W/m²) ====== Shortwave radiation(SW)
LW_IN_F - Incoming longwave radiation, LWDOWN (W/m²) ======== Longwave radiation(LW)
VPD_F - Vapor pressure deficit, VPD (hPa) ============== Vapor Pressure Deficit(VPD)
RH - Relative humidity(RH)
Atmospheric pressure(PA)
P_F - Rainfall, RAIN (mm) ======================================== Precipitation(P)
WS_F - Windspeed, WIND (m/s) ======================================== Wind speed(WS)
WD - Wind direction(WD)
PPFD_IN - Photosynthetic photon flux density (PPFD)
TS_F_MDS_# - Soil temperature, TSOIL (°C) =============== Soil temperature(TS) 1最浅

ENERGY PROCESSING
NETRAD - Net radiation, RN (W/m²) ============================ Net radiation(NETRAD)
G_F_MDS - Soil heat flux, GF (W/m²) ============================ Soil heat flux(G_F)
LE_F_MDS - Latent heat flux QLE (W/m²) ====================== Latent heat flux(LE_F)
H_F_MDS - Sensible heat flux QH (W/m²) ===================== Sensible heat flux(H_F)

NET ECOSYSTEM EXCHANGE 
NEE_VUT_REF - Net Ecosystem Exchange, NEE (gC/m²/d) ================= Net Ecosystem Exchange(NEE_VUT_REF)
GPP_NT_VUT_REF/GPP_DT_VUT_REF - Gross primary production, GPP (gC/m²/d)  ================= Gross Primary Production(GPP_NT_VUT_REF)
RECO_NT_VUT_REF/RECO_DT_VUT_REF - Ecosystem respiration, Reco (gC/m²/d)  ================= Ecosystem Respiration(RECO_NT_VUT_REF)
注意生态系统GPP和呼吸有日间和夜间两种计算方法
```{r Load your flux_selected}
flux_guy <- read.csv('/Users/quan/projects/Trait_Orchidee/Flux/Fluxdata/FLX_GF-Guy_FLUXNET2015_FULLSET_2004-2020_Evergreen_Broadleaf_Forests/FLX_GF-Guy_FLUXNET2015_FULLSET_DD_2004-2020_beta-3.csv')
flux_Bra <- read.csv('/Users/quan/projects/Trait_Orchidee/Flux/Fluxdata/FLX_BE-Bra_FLUXNET2015_FULLSET_1996-2020_Evergreen_Needleleaf_Forests/FLX_BE-Bra_FLUXNET2015_FULLSET_DD_1996-2020_beta-3.csv')

vars_short <- c("Time","T","T_NIGHT","T_DAY","SWDOWN","LWDOWN",
                "VPD","RAIN","WIND","RN","TSOIL","GF","QLE","QH",
                "NEE","RECO","GPP") 
vars_long <- c("TIMESTAMP","TA_F","TA_F_NIGHT","TA_F_DAY","SW_IN_F","LW_IN_F",
               "VPD_F","P_F","WS_F",'NETRAD','TS_F_MDS_1','G_F_MDS','LE_F_MDS','H_F_MDS',
               'NEE_VUT_REF','RECO_DT_VUT_REF','GPP_DT_VUT_REF')
vars_long[!(vars_long %in% colnames(flux_Bra))]
# sort(colnames(flux_Bra))
flux_selected <- flux_Bra[vars_long]
colnames(flux_selected) <- vars_short
```

**Additional helpful Metaflux_selected:**

- World map of Köppen-Geiger climate classification (PDF)
- Site information: Find from [fluxnet sites list](https://fluxnet.org/sites/site-list-and-pages) or just google the name of your site

- Check and filter missing (NA) or impossible (e.g. very larges, -9999) values
```{r}
## Step 1
# Filter weird values
vars <- c("T","T_NIGHT","T_DAY","SWDOWN","LWDOWN" ,"VPD","RAIN","WIND","RN","TSOIL","GF","QLE","QH","NEE","RECO","GPP") 
for (var in vars){
  flux_selected[[var]][abs(flux_selected[[var]])>1e20] <- NA
  flux_selected[[var]][(flux_selected[[var]])< -5000] <- NA
}


```

- Convert variables into easy interpretable units (land fluxes into kgC /m²/year or tC /ha/year)
Net Ecosystem Exchange, NEE (gC/m²/d); Gross primary production, GPP (gC/m²/d); Ecosystem respiration, Reco (gC/m²/d)
```{r}
# Conversions
flux_selected.converted <- flux_selected %>% 
  mutate(Time = as.Date(as.character(Time),"%Y%m%d"),
          year = year(Time),month= month(Time),yday = yday(Time)) 

flux_selected.converted$GPP <- flux_selected.converted$GPP * 365 * 10000/1000000
flux_selected.converted$NEE <- flux_selected.converted$NEE * 365 * 10000/1000000
flux_selected.converted$RECO <- flux_selected.converted$RECO * 365 * 10000/1000000
```

- Calculate the minimal and maximal value for each variable, as a quick check
```{r}
# Min and max values
flux_selected.converted %>% dplyr::select(-c("Time","year","yday")) %>% summarise_all(min, na.rm = TRUE)
flux_selected.converted %>% dplyr::select(-c("Time","year","yday")) %>% summarise_all(max, na.rm = TRUE)
```

- Calculate annual sum of precipitation, and the annual mean for each other variable. 
```{r}
# Sum for Precip, mean for the rest
flux_selected.converted %>% dplyr::select(c("year","RAIN","yday")) %>% group_by(year) %>% summarise(Precip = sum(RAIN, na.rm = TRUE))
flux_selected.converted %>% dplyr::select(-c("RAIN","Time","yday")) %>% group_by(year) %>% summarise_all(mean, na.rm = TRUE)

flux_selected.converted %>% dplyr::select(c("year","RAIN","yday")) %>% group_by(year) %>% summarise(Precip = sum(RAIN)) %>% pull(Precip) %>% mean(na.rm = TRUE)
flux_selected.converted %>% dplyr::select(-c("RAIN","Time","yday","year")) %>% summarise_all(mean, na.rm = TRUE)

```


### Step 2: Mean annual cycles
Seasonal cycle, patterns, extreme events 

```{r}
flux_selected.cycle <- flux_selected.converted %>% dplyr::select(-c("year","Time")) %>% group_by(yday) %>% summarise_all(mean,na.rm = TRUE)
flux_selected.cycle.long <- flux_selected.cycle %>% pivot_longer(cols = -c("yday"),
                                               names_to = "var",
                                               values_to = "value")

flux_selected.cycle.long.years <- flux_selected.converted %>% dplyr::select(-c("Time"))%>% group_by(year,yday) %>% summarise_all(mean,na.rm = TRUE) %>% pivot_longer(cols = -c("yday","year"),
                                                                                               names_to = "var",
                                                                                               values_to = "value")
```

- **Plot 1**: Produce mean seasonal cycle graphs (= you compile your flux_selected to obtain one average year) for each meteorological variable (in a single plot)
- Check the correspondence of these plots: do patterns concur or differ? Why? Try to explain the causal relationship. 
```{r  Plot 1}
# Plot 1
ggplot(data = flux_selected.cycle.long %>% 
         dplyr::filter(var %in%  c("T","T_NIGHT","T_DAY","SWDOWN","LWDOWN" ,"VPD","RAIN","WIND","RN","TSOIL"))) +
  geom_line(aes(x = yday,y = value)) +
    geom_smooth(aes(x = yday,y = value),colour="red", size=1, method="loess",
              span=1/5, se=FALSE) +
  facet_wrap(~var,scales = "free") +

  theme_bw()

```

- **Plot 2**: Compare your average year with every single year of flux_selected. Are there any extreme events in your flux_selected? Document it.
```{r  Plot 2}
# Plot 2 
ggplot(data = flux_selected.cycle.long %>% dplyr::filter(var %in% c("T","T_NIGHT","T_DAY","SWDOWN","LWDOWN" ,"VPD","RAIN","WIND","RN","TSOIL"))) +
  geom_line(data = flux_selected.cycle.long.years  %>% dplyr::filter(var %in% c("T","T_NIGHT","T_DAY","SWDOWN","LWDOWN" ,"VPD","RAIN","WIND","RN","TSOIL")),
            aes(x = yday,y = value,group = as.factor(year)),color = "grey") +
  geom_line(aes(x = yday,y = value)) +
  facet_wrap(~var,scales = "free") +
  theme_bw()
```

- **Plot 3**: Using bar plots, plot the average precipitation yearly cycle
```{r  Plot 3}
ggplot(flux_selected.cycle,aes(x=yday,y=RAIN)) +
   geom_col() +
  scale_x_continuous(breaks = seq(1,365,30)) +
  theme_bw()
```


### Step 3: Surface energy balance
Radiation balance, energy closure, Bowen ratio 

- **Plot 4**: Plot mean annual cycles for Rn, Qh, Qle, GF (W/m²), in one graph  and interpret the annual cycles.
```{r }
# plot 4
ggplot(data = flux_selected.cycle.long %>% dplyr::filter(var %in% c("QLE","QH","RN","GF"))) +
  geom_line(aes(x = yday,y = value,color = as.factor(var))) +
  theme_bw()
```

- **Plot 5**: Scatterplot of energy balance (Qh+Qle vs. Rn-GF) (W/m²), select only 1 year of your flux_selectedset for this plot, fit a linear regression line (include equation + R²). Compare with 1:1 line.
```{r}
# Plot 5 
flux_selected.balance <- flux_selected.cycle.long %>% 
  filter(var %in% c("QLE","QH","RN","GF")) %>% pivot_wider(names_from = var,values_from = value) %>% 
  mutate(y = QLE + QH, x = RN - GF)

ggplot(data = flux_selected.balance,
       aes(x = x,y = y)) +
  geom_point() +
  geom_abline(slope = 1,color = "black",linetype = 2) +
  labs(x = "Rn - G (W/m²)",
       y = "Qle + Qh (W/m²)") +
  geom_smooth(method='lm', formula= y~x,se = FALSE,color = "red") + 
  theme_bw()

LM <- lm(y ~ x, data = flux_selected.balance,)
coef(LM)
summary((LM))$r.squared
```
- Analyze the closure of the energy balance. How much is the energy gap? What causes this? Did you expect such magnitude? Refer to literature in your answer.
```{r}
hist(flux_selected.balance$QH/flux_selected.balance$QLE)
```

- Calculate the Bowen ratio (= Qh/Qle).
```{r}
(flux_selected.balance$QH/flux_selected.balance$QLE) %>% mean()
```


### Step 4: Carbon fluxes
Carbon balance, carbon closure

- **Plot 6**: mean annual cycles for GPP, Reco and NEE, all in one plot
```{r}
ggplot(data = flux_selected.cycle.long %>% dplyr::filter(var %in% c("GPP","NEE","RECO"))) +
  geom_line(aes(x = yday,y = value,color = as.factor(var))) +
  theme_bw()
```

- Interpret the annual cycles. Is the site more a carbon source (NEE > 0) or a carbon sink (NEE < 0)?
```{r}
ggplot(data = flux_selected.cycle %>% mutate(balance = GPP - RECO - NEE)) +
  geom_histogram(aes(x = balance)) +
  theme_bw()

ggplot(data = flux_selected.converted,
       aes(x = GPP - RECO,y = -NEE,color = as.factor(year))) +
  geom_point() +
  geom_abline(slope = 1,color = "black",linetype = 2) +
  geom_smooth(method='lm', formula= y~x,se = FALSE,color = "red") + 
  theme_bw()

coef(lm(data = flux_selected.converted %>% mutate(NEP = - GPP + RECO),
        formula =  NEE ~ NEP))

```


### Step 5: Interpretation
Climate type, links between variables, relation with ecosystem 

- **Plot 7**: Are there links between different meteorological variables?
```{r}
flux_selected_corr <- flux_selected.converted %>% dplyr::select(-c("Time","year","yday"))
M <- cor(flux_selected_corr,use="complete.obs")
corrplot(M, method="pie",type = "upper")

p.mat <- cor.mtest(flux_selected_corr)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, method="color",  
         type="upper",
         addCoef.col = "black",
         tl.col="black", tl.srt=45,
         diag=FALSE)
```

- **Plot 8**: Are some of the land fluxes (GPP, Reco, NEE) correlated with one or several meteorological variables?
- Compare climate type with that on other locations (e.g. other groups, textbook examples, papers, ...)
- Comment on extreme events, inter-annual variability, anomalies, …
- What is the relation between the climate and the site characteristics? What is the role of vegetation in this climate system?
- Any surprising details?
```{r}
# Plot 8
flux_selected_corr2 <- flux_selected.cycle %>% dplyr::select(-c("yday")) %>% dplyr::select(c("GPP","RECO","NEE","T_DAY","SWDOWN"))
pairs.panels(flux_selected_corr2, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = FALSE )

```



### Useful R cheatsheets
1. Summary of Rstudio functions: https://raw.githubusercontent.com/rstudio/cheatsheets/master/rstudio-ide.pdf
2. R basic commands: https://raw.githubusercontent.com/rstudio/cheatsheets/master/base-r.pdf
3. flux_selected visualization: https://raw.githubusercontent.com/rstudio/cheatsheets/master/flux_selected-visualization-2.1.pdf
4. Work with dates: https://raw.githubusercontent.com/rstudio/cheatsheets/main/lubridate.pdf
5. More R cheatsheets: https://rstudio.com/resources/cheatsheets/


